Self Referencing Sequences

### 2014-10-28 draft.tex
Edited Density

In this section, we analyze two things: the first is how to take the rule based methods of sequence generation to show that if a density exists, we can find what it is. We will explain the method and give an example applied to the case above of $G = \{ 1, 3\}$ and $G = \{ 2, 4 \}$. In the second part, we will show an ad hoc way to bound the density of $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally intractable for smaller bounds since the number of cases to check increases exponentially.

\subsection{Determining the density}

We have found a technique to compute the density of numbers in a self-referential sequence (if one exists) for some generating sets. In particular, suppose we have a generating set $G$ with a set of rules $R$ that satisfy the conditions from the section above. For each rule $r \in R$, we will say denote the rule as $r_l \rightarrow r_r$ as the inputs and the outputs.

@@

For example, lets look at the rules we have determined for $G = \{ 1, 3 \}$. 
\begin{align*}
\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
\text{Rule }B: 3,1 &\to 3,3,3,1\\
\text{Rule }C: 1,1 &\to 3,1
\end{align*}
Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurences of $A$ in $s^{(i)}$, then 
\[ A_i = B_{i-1} + A_{i-1} \]
Likewise, we have
\[ B_i = A_{i-1} + B_{i-1} + C_{i-1} \]
\[ C_i = A_{i-1} \]
So now we can take the fraction by dividing by the total number of them:
\[ A_i' = \dfrac{B_{i-1}' + A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
\[ B_i' = \dfrac{A_{i-1}' + B_{i-1}' + C_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
\[ C_i' = \dfrac{A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearence of each and get the density of the numbers.

In particular, the density of 3's is approxinately $0.602$ by computing the solution to the set of quadratic equations.

\subsection{Density bounds when $G = \{1, 2\}$}

As shown in the previous section, we cannot establish the set of rules we want when $G = \{1 , 2\}$; however, there is some structure that we can exploit. In particular, we know that the sequence $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence. This is because they would be part of a block of size at least $3$.

@@

In general, this gives a procedure for computing the limiting density. For example, now we know that there must be at least one $2$ for every three numbers. So now we can look at how many $2$'s are in any given string of length $9$. We know there must be at least $3$, but each consecutive $3$ elements must contain a $3$. There are only two cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
\[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
\[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
In each of these cases, we can read the subsequence that must have generated them part of them. In order, they are
\[ 2, 1, 2, 1, 2 \]
\[ 1, 2, 1, 2, 1\]
\[ 2, 1, 2, 1, 2 \]
And these are generated by $1, 1, 1$. Which means that none of these are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9. Therefore, we have shown that the density must be bounded by
\[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
In general, we can continue this process, and we did, to show that
\[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
Of course, the number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on. Its not hard to see that this number grows exponentially fast, making it computationally difficult to analyze more cases.

We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, generalizing the procedure to improve the bound seems dificult.   
\end{document}

>>>>

A topic of interest when considering a self referencing sequence is the relative density of the various elements of the generating set. It is unclear if such a density even exists and if so how to compute it for arbitrary generating sets. In this section, we show how to use substitution rules to compute relative densities for certain sequences if we assume such densities exists. We explain the method and compute the relative density of 3's for $G = \{ 1, 3\}$. We will also produce bounds on the density of $2$'s $G = \{ 1, 2\}$ in a way that suggests that the density converges to 2; however, this method is computationally difficult for smaller bounds since the number of cases to check increases exponentially.

To begin we will discuss a technique that can be used to compute the densities for some generating sets. In particular, suppose we have a generating set $G$ with a set of rules $R$ that satisfy the conditions from the section above. For each rule $r \in R$, we will say denote the rule as $r_l \rightarrow r_r$ as the inputs and the outputs.

@@

For example, lets look at the rules we have determined for $G = \{ 1, 3 \}$. 
\begin{align*}
\text{Rule }A: 3,3 &\to 3,3,3,1,1,1 \\
\text{Rule }B: 3,1 &\to 3,3,3,1\\
\text{Rule }C: 1,1 &\to 3,1
\end{align*}
Then we can let $3,3,3,1,1,1 = A$, $3,3,3,1 = B$ and $3, 1 = C$. If $A_i$ is the number of occurrences of $A$ in $s^{(i)}$, then 
\[ A_i = B_{i-1} + A_{i-1} \]
Likewise, we have
\[ B_i = A_{i-1} + B_{i-1} + C_{i-1} \]
\[ C_i = A_{i-1} \]
So now we can take the fraction by dividing by the total number of them:
\[ A_i' = \dfrac{B_{i-1}' + A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
\[ B_i' = \dfrac{A_{i-1}' + B_{i-1}' + C_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
\[ C_i' = \dfrac{A_{i-1}'}{3A_{i-1}' + 2B_{i-1}' + C_{i-1}'} \]
So if we assume a density, there will be a stable state which will be the density, so in the case where $A_i'=A_{i-1}', B_i' = B_{i-1}', C_i'=C_{i-1}'$, we can solve for the fractional appearance of each and get the density of the numbers.

We can compute solutions to this set of quadratic equations and find the density of 3's is approximately $0.602$.

As discussed in the previous section, it is unclear how to establish a set of rules for $G = \{1 , 2\}$; however, there is some structure that we can exploit. In particular, we know that the sequences $1, 1, 1$ and $2, 2, 2$ will never appear in the self-referential sequence. This is because such sequences would be part of a block of size at least $3$.

@@

In general, this gives a procedure for computing the limiting density. For example, now we know that there must be at least one $2$ for every three numbers. So now we can look at how many $2$'s are in any given string of length $9$. We know there must be at least three $2$'s. There are only three cases where there are three 2's in a segment of $9$ numbers in the sequence, namely
\[ 2, 1, 1, 2, 1, 1 ,2 , 1, 1 \]
\[ 1, 2, 1, 1, 2, 1, 1, 2, 1 \]
\[ 1, 1, 2, 1, 1, 2, 1, 1, 2 \]
In each of these cases, we can read the subsequence that must have generated them part of them. In order, they are
\[ 2, 1, 2, 1, 2 \]
\[ 1, 2, 1, 2, 1\]
\[ 2, 1, 2, 1, 2 \]
but each of these are generated by $1, 1, 1$. This means that none of the sequences of length 9 are possible, so in addition to having one 2 for every group of 3, there is another 2 for every group of 9. Therefore, we have shown that the density must be bounded by
\[ \dfrac{1}{3} + \dfrac{1}{9} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} \]
In general, we can continue this process, and we did, to show that
\[ \dfrac{1}{3} + \dfrac{1}{9} + \dfrac{1}{27} \leq d_2 \leq 1 - \dfrac{1}{3} - \dfrac{1}{9} - \dfrac{1}{27} \]
Of course, the number of cases to check increases very quickly: there are $3$ choices to put the $2$ in the groups of $3$, $3$ choices to put the 3 groups of 2's and 2 positions within these groups for the groups of 9, and so on. Its not hard to see that this number grows exponentially fast, making it computationally difficult to analyze more cases.

We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, generalizing the procedure to improve the bound seems difficult.   
\end{document}
